### 2025-08-04 paper/camera/sects/conclusion.tex
to 7 pages for contribution

Despite these encouraging results, several avenues remain for future exploration. One key direction is scaling beyond the current limit of tens of agents. Another challenge lies in environments with extreme partial observability, such as the discovery scenario, where limited sensory input hinders learning. Expanding the evaluation to a broader set of environments—particularly those from the cooperative AI community~\cite{DBLP:journals/corr/abs-2211-13746} and adversarial settings—would further validate the robustness and versatility of \fname{}.

Addressing the cold-start problem more effectively could also enhance performance, particularly in environments with sparse or delayed rewards. Several promising techniques for mitigating this issue include:

\begin{itemize}
    \item \textbf{World model pre-training:} Leveraging datasets generated by alternative algorithms or human demonstrations to initialize the world model before policy training begins.
    \item \textbf{Model Predictive Path Integral (MPPI) control~\cite{Alvarez24}:} Using MPPI to generate high-quality initial trajectories that can seed the world model and guide early exploration.
\end{itemize}

The latter approach may also be integrated into \fname{} during training to periodically introduce diverse trajectories, potentially improving robustness and convergence speed.

>>>>

Despite these encouraging results, several avenues remain for future exploration: scaling beyond tens of agents, handling extreme partial observability (e.g., discovery scenarios with limited sensory input), and expanding evaluation to broader environments---particularly from the cooperative AI community~\cite{DBLP:journals/corr/abs-2211-13746} and adversarial settings—to further validate \fname{}'s robustness and versatility.

Addressing the cold-start problem more effectively could also enhance performance, particularly in environments with sparse or delayed rewards. Several promising techniques for mitigating this issue include:
\begin{enumerate*}[label=\emph{\roman*})]
    \item leveraging datasets generated by alternative algorithms or human demonstrations to initialize the world model before policy training begins (world-model pre-training).
    \item using Model Predictive Path Integral (MPPI)~\cite{Alvarez24}  to generate high-quality initial trajectories that can seed the world model and guide early exploration.
\end{enumerate*}
%
The latter approach may also be integrated into \fname{} during training to periodically introduce diverse trajectories, potentially improving robustness and convergence speed.

### 2025-08-04 paper/camera/sects/experiments.tex
to 7 pages for contribution

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{figs/grads-transformer-transport.pdf}
    \caption{Policy gradient norms in the Transport environment over 5000 epochs. Spikes in gradient norm for SHAC (epoch 2700) and \fname{} (epoch 3200) correspond to generalization events when agents begin frequent interactions with the package and each other.}\label{fig:grads-transformer-transport}
    \vspace{0.5cm}
\end{figure}

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.9\textwidth]{figs/main-transformer.pdf}
    \caption{Performance comparison of \fname{}, PPO/MAPPO, and SHAC across different scenarios (Dispersion, Transport, Discovery, and Sampling) using the Transformer architecture for multi-agent settings and the MLP architecture for single-agent settings. The results show the mean and standard deviation of rewards over 3 runs. For results using the MLP architecture in all scenarios, refer to \Cref{apx:fig:experiments-mlp}.}%
    \label{fig:experiments}
\end{figure*}

\begin{table*}
    \centering
    \input tables/max.tex
    \vspace{0.2cm}
    \caption{Normalized rewards (relative to the best performing model) for the different scenarios. Best results are in bold.}\label{tab:max-rewards}
\end{table*}

>>>>

\begin{figure}
    \centering
    \includegraphics[width=0.7\columnwidth]{figs/grads-transformer-transport.pdf}
    \caption{Policy gradient norms in the Transport environment over 5000 epochs. Spikes in gradient norm for SHAC (epoch 2700) and \fname{} (epoch 3200) correspond to generalization events when agents begin frequent interactions with the package and each other.}\label{fig:grads-transformer-transport}
    \vspace{0.5cm}
\end{figure}

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=0.75\textwidth]{figs/main-transformer.pdf}
    \caption{Performance comparison of \fname{}, PPO/MAPPO, and SHAC across different scenarios (Dispersion, Transport, Discovery, and Sampling) using the Transformer architecture for multi-agent settings and the MLP architecture for single-agent settings. The results show the mean and standard deviation of rewards over 3 runs. For results using the MLP architecture in all scenarios, refer to \Cref{apx:fig:experiments-mlp}.}%
    \label{fig:experiments}
\end{figure*}

\begin{table}[t]
    \centering
    \input tables/max.tex
    \vspace{0.2cm}
    \caption{Normalized rewards (relative to the best performing model) for the different scenarios. Best results are in bold. \textsc{D} stands for Dispersion, \textsc{T} for Transport, \textsc{Di} for Discovery, and \textsc{S} for Sampling.}\label{tab:max-rewards}.
\end{table}

### 2025-08-04 paper/camera/sects/limitations.tex
to 7 pages for contribution

A central challenge lies in learning the world model, which depends on rich, meaningful interactions between agents and the environment. These interactions often yield sparse or delayed rewards, leading to a cold-start problem: until the world model is sufficiently trained, agents may struggle to explore or act effectively, even when guided by stochastic policies. This issue can be partially addressed by seeding the learning process with a cold-start dataset—either manually designed or generated using alternative algorithms such as PPO or MPPI~\cite{Alvarez24}. Such data helps bootstrap the world model, providing a more stable foundation for training.

Despite careful attention to implementation, including reference to best practices (e.g., the well-known "37 Implementation Details of Proximal Policy Optimization" blog~\cite{Shengyi22}), we observed that PPO and MAPPO performed suboptimally in our experiments. To ensure the validity of these findings, we cross-verified results using a third-party implementation~\cite{Bettini24}, which yielded consistent outcomes.

Although early stopping is primarily used to conserve computational resources, it can give the false impression that the algorithm is inherently robust to policy collapse. In reality, we observed multiple instances of partial policy collapse—some of which recovered—and a few cases of complete collapse without recovery. One example of partial collapse appears in the five-agent variant of the dispersion scenario shown in \Cref{apx:fig:experiments-mlp}.

The discovery scenario itself poses considerable difficulty for all tested algorithms due to its inherent partial observability. Agents rely solely on a simulated LiDAR system for perception, which provides a narrow, local view of the environment. This makes it particularly difficult for any method—including \fname{}—to construct a complete and accurate world model, often resulting in suboptimal performance. Ultimately, this may indicate that partial observability may be challenging to overcome with the current approach.

\fname{} addresses the limitations of SHAC by introducing additional training components, specifically the transition and reward networks. However, this improvement comes at the cost of increased computational overhead. The training burden is especially pronounced in environments with a large number of agents, where the transition network must capture complex inter-agent dynamics. While this trade-off is acceptable in many scenarios, it significantly slows down the training process and demands computational resources that may not be readily available in all settings.

Regarding interpretability, \fname{} does not offer significant advantages over its predecessors. As the policy is governed by a neural network, understanding or explaining its behavior remains a challenging task. This limitation is inherent to most deep reinforcement learning methods and highlights the ongoing need for techniques that enhance policy transparency and interpretability.

>>>>

\paragraph{World Model Learning and Cold-Start Problem}
A central challenge lies in learning the world model, which depends on rich agent-environment interactions. Sparse or delayed rewards create a cold-start problem: until the world model is sufficiently trained, agents struggle to explore effectively. This can be addressed by seeding with a cold-start dataset---manually designed or generated using algorithms like PPO or MPPI~\cite{Alvarez24}---to bootstrap the world model and provide a stable training foundation.

\paragraph{Baseline Performance and Implementation}
Despite careful attention to implementation, including reference to best practices (e.g., the well-known ``37 Implementation Details of Proximal Policy Optimization'' blog~\cite{Shengyi22}), we observed that PPO and MAPPO performed suboptimally in our experiments. To ensure the validity of these findings, we cross-verified results using a third-party implementation~\cite{Bettini24}, which yielded consistent outcomes.

\paragraph{Policy Collapse and Robustness}
Although early stopping is primarily used to conserve computational resources, it can give the false impression that the algorithm is inherently robust to policy collapse. In reality, we observed multiple instances of partial policy collapse---some of which recovered---and a few cases of complete collapse without recovery. One example of partial collapse appears in the five-agent variant of the dispersion scenario shown in \Cref{apx:fig:experiments-mlp}.

\paragraph{Partial Observability Challenges}
The discovery scenario poses considerable difficulty for all algorithms due to inherent partial observability. Agents rely solely on simulated LiDAR for perception, providing only a narrow, local view. This makes it difficult for any method---including \fname{}---to construct accurate world models, often resulting in suboptimal performance and suggesting that partial observability remains challenging for the current approach.

\paragraph{Computational Overhead}
\fname{} addresses SHAC's limitations by adding transition and reward networks, but this increases computational overhead. Training burden is especially pronounced with many agents, where the transition network must capture complex inter-agent dynamics. While this trade-off is often acceptable, it significantly slows training and demands substantial computational resources.

\paragraph{Interpretability}
Regarding interpretability, \fname{} does not offer significant advantages over its predecessors. As the policy is governed by a neural network, understanding or explaining its behavior remains a challenging task. This limitation is inherent to most deep reinforcement learning methods and highlights the ongoing need for techniques that enhance policy transparency and interpretability.
